---
title: "Cole's Code"
output: html_document
date: "2024-06-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(broom)
library(hockeyR)
```



## Thoughts on measuring consistency


- Measure the difference of playing statistics season-over-season
- Weight the metrics differently based on importance
  - consistency score will have more consistent players near 0 & inconsistent players much greater than 0
- playing metrics will have to be per game or per 60 measures for unbiased information


- Use pivot_wider function to increase number of columns in order to take the differences between playing metrics

- Use the average of differences to measure consistency



```{r}
basic_2023_24 <- read_csv("../Data/Basic Stats/basic_stats_23_24.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2023)

basic_2022_23 <- read_csv("../Data/Basic Stats/basic_stats_22_23.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2022)

basic_2021_22 <- read_csv("../Data/Basic Stats/basic_stats_21_22.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2021)

basic_2020_21 <- read_csv("../Data/Basic Stats/basic_stats_20_21.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2020)

basic_2019_20 <- read_csv("../Data/Basic Stats/basic_stats_19_20.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2019)

basic_2018_19 <- read_csv("../Data/Basic Stats/basic_stats_18_19.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2018)

basic_2017_18 <- read_csv("../Data/Basic Stats/basic_stats_17_18.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2017)

basic_2016_17 <- read_csv("../Data/Basic Stats/basic_stats_16_17.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2016)
```



```{r}
basic_df <- rbind(basic_2016_17, basic_2017_18, basic_2018_19, basic_2019_20, basic_2020_21, basic_2021_22, basic_2022_23, basic_2023_24)
```



```{r}
long_basic_df <- basic_df |> 
  pivot_wider(id_cols = c(Player, PlayerID), values_from = c(G, A, PTS, `+/-`, `S%`, PS), names_from = start_year)

long_basic_df <- long_basic_df %>% 
  unnest() %>% 
  replace(. == "NULL", NA)
  # mutate(across(G_2023:PS_2016, ~replace_na(.x, 0)))

# long_basic_df_exp <- long_basic_df |> 
#   mutate(Gdiff_23_22 = abs(G_2023 - G_2022),
#          Gdiff_22_21 = abs(G_2022 - G_2021),
#          Gdiff_21_20 = abs(G_2021 - G_2020),
#          Gdiff_20_19 = abs(G_2020 - G_2019),
#          Gdiff_19_18 = abs(G_2019 - G_2018),
#          Gdiff_18_17 = abs(G_2018 - G_2017),
#          Gdiff_17_16 = abs(G_2017 - G_2016)
#          ) |>
#   group_by(Player, PlayerID) |>
#   summarise(total_Gdiff = sum(Gdiff_23_22, Gdiff_22_21, Gdiff_21_20, Gdiff_20_19, Gdiff_19_18, Gdiff_18_17, Gdiff_17_16))
```




```{r}
advanced_2023_24 <- read_csv("../Data/Advanced Stats/advanced_stats_23_24.csv") |> 
  mutate(start_year = 2023)

advanced_2022_23 <- read_csv("../Data/Advanced Stats/advanced_stats_22_23.csv") |> 
  mutate(start_year = 2022)

advanced_2021_22 <- read_csv("../Data/Advanced Stats/advanced_stats_21_22.csv") |> 
  mutate(start_year = 2021)

advanced_2020_21 <- read_csv("../Data/Advanced Stats/advanced_stats_20_21.csv") |> 
  mutate(start_year = 2020)

advanced_2019_20 <- read_csv("../Data/Advanced Stats/advanced_stats_19_20.csv") |> 
  mutate(start_year = 2019)

advanced_2018_19 <- read_csv("../Data/Advanced Stats/advanced_stats_18_19.csv") |> 
  mutate(start_year = 2018)

advanced_2017_18 <- read_csv("../Data/Advanced Stats/advanced_stats_17_18.csv") |>
  mutate(start_year = 2017)

advanced_2016_17 <- read_csv("../Data/Advanced Stats/advanced_stats_16_17.csv") |>
  mutate(start_year = 2016)

advanced_df <- rbind(advanced_2016_17, advanced_2017_18, advanced_2018_19, advanced_2019_20, advanced_2020_21, advanced_2021_22, advanced_2022_23, advanced_2023_24)

advanced_df <- advanced_df |> filter(GP >= 20) |> select(Player, PlayerID, GP, `oiSH%`, `oiSV%`, PDO, TK, GV, `CF%`, `FF%`)
```




```{r}
long_adv_df <- advanced_df_df |> 
  filter(Pos != "G", GP >= 20) |> 
  pivot_wider(id_cols = c(Player), values_from = c(GP, `oiSH%`, `oiSV%`, PDO, TK, GV, `CF%`, `FF%`), names_from = start_year)

long_adv_df <- long_adv_df %>% 
  unnest() %>% 
  replace(. == "NULL", NA)
```




```{r}
misc_2023_24 <- read_csv("../Data/Misc Stats/misc_stats_23_24.csv") |> 
  mutate(start_year = 2023)

misc_2022_23 <- read_csv("../Data/Misc Stats/misc_stats_22_23.csv") |>
  mutate(start_year = 2022)

misc_2021_22 <- read_csv("../Data/Misc Stats/misc_stats_21_22.csv") |> 
  mutate(start_year = 2021)

misc_2020_21 <- read_csv("../Data/Misc Stats/misc_stats_20_21.csv") |> 
  mutate(start_year = 2020)

misc_2019_20 <- read_csv("../Data/Misc Stats/misc_stats_19_20.csv") |> 
  mutate(start_year = 2019)

misc_2018_19 <- read_csv("../Data/Misc Stats/misc_stats_18_19.csv") |> 
  mutate(start_year = 2018)

misc_2017_18 <- read_csv("../Data/Misc Stats/misc_stats_17_18.csv") |> 
  mutate(start_year = 2017)

misc_2016_17 <- read_csv("../Data/Misc Stats/misc_stats_16_17.csv") |> 
  mutate(start_year = 2016)

misc_df <- rbind(misc_2016_17, misc_2017_18, misc_2018_19, misc_2019_20, misc_2020_21, misc_2021_22, misc_2022_23, misc_2023_24)

misc_df <- misc_df |> 
  mutate(Pos = str_extract(Pos, "[A-Z]"))
```



```{r}
num_seasons <- misc_df |> 
  filter(Pos != "G") |> 
  group_by(Player) |> 
  count() |> 
  ungroup() |> 
  mutate(num_seasons = n) |> 
  unique()

player_pos <- misc_df |> 
  filter(Pos != "G") |> 
  group_by(Player) |> 
  summarise(Pos = ifelse(Pos %in% c("C", "L", "R", "W", "F"), "F", "D")) |> 
  unique()
```



```{r}
long_misc_df <- misc_df |> 
  filter(Pos != "G", GP >= 20) |> 
  pivot_wider(id_cols = c(Player), values_from = c(`G/GP`, `A/GP`, `PTS/GP`, `GC/GP`, `S/GP`, `+/-`, `OPS`, `DPS`, `PS`), names_from = start_year)

long_misc_df <- long_misc_df %>% 
  unnest() %>% 
  replace(. == "NULL", NA)
  # mutate(across(`G/GP_2023`:`PS_2016`, ~replace_na(.x, 0)))

long_misc_df_exp <- long_misc_df |>
  group_by(Player) |> 
  summarise(ptspg_diff_23_22 = abs(`PTS/GP_2023` - `PTS/GP_2022`),
         ptspg_diff_22_21 = abs(`PTS/GP_2022` - `PTS/GP_2021`),
         ptspg_diff_21_20 = abs(`PTS/GP_2021` - `PTS/GP_2020`),
         ptspg_diff_20_19 = abs(`PTS/GP_2020` - `PTS/GP_2019`),
         ptspg_diff_19_18 = abs(`PTS/GP_2019` - `PTS/GP_2018`),
         ptspg_diff_18_17 = abs(`PTS/GP_2018` - `PTS/GP_2017`),
         ptspg_diff_17_16 = abs(`PTS/GP_2017` - `PTS/GP_2016`),
         
         ops_diff_23_22 = abs(OPS_2023 - OPS_2022),
         ops_diff_22_21 = abs(OPS_2022 - OPS_2021),
         ops_diff_21_20 = abs(OPS_2021 - OPS_2020),
         ops_diff_20_19 = abs(OPS_2020 - OPS_2019),
         ops_diff_19_18 = abs(OPS_2019 - OPS_2018),
         ops_diff_18_17 = abs(OPS_2018 - OPS_2017),
         ops_diff_17_16 = abs(OPS_2017 - OPS_2016),
         
         p_m_diff_23_22 = abs(`+/-_2023` - `+/-_2022`),
         p_m_diff_22_21 = abs(`+/-_2022` - `+/-_2021`),
         p_m_diff_21_20 = abs(`+/-_2021` - `+/-_2020`),
         p_m_diff_20_19 = abs(`+/-_2020` - `+/-_2019`),
         p_m_diff_19_18 = abs(`+/-_2019` - `+/-_2018`),
         p_m_diff_18_17 = abs(`+/-_2018` - `+/-_2017`),
         p_m_diff_17_16 = abs(`+/-_2017` - `+/-_2016`),
         
         total_ptspg_diff = sum(ptspg_diff_23_22, ptspg_diff_22_21, ptspg_diff_21_20, ptspg_diff_20_19, ptspg_diff_19_18, ptspg_diff_18_17, ptspg_diff_17_16, na.rm = TRUE),
         
         total_ops_diff = sum(ops_diff_23_22, ops_diff_22_21, ops_diff_21_20, ops_diff_20_19, ops_diff_19_18, ops_diff_18_17, ops_diff_17_16, na.rm = TRUE),
         
         total_p_m_diff = sum(p_m_diff_23_22, p_m_diff_22_21, p_m_diff_21_20, p_m_diff_20_19, p_m_diff_19_18, p_m_diff_18_17, p_m_diff_17_16, na.rm = TRUE),
         
         # std_ptspg_diff = abs(log(total_ptspg_diff)),
         
         # std_ops_diff = log(total_ops_diff),
         
         # std_p_m_diff = log(total_p_m_diff),
         
            consistency_score = sum(0.5 * total_ptspg_diff, 0.2 * total_ops_diff, 0.3 * total_p_m_diff)
  )


long_misc_df_exp <- long_misc_df_exp |> full_join(player_pos)

long_misc_df_exp <- long_misc_df_exp |> full_join(num_seasons) |> select(-n)

long_misc_df_exp <- long_misc_df_exp |> 
  mutate(consistency_score = round(consistency_score / num_seasons, 3))
```



```{r}
long_misc_df_exp |> 
  group_by(Player) |> 
  arrange(desc(total_ptspg_diff)) |> 
  head(n = 10) |> 
  ggplot(aes(y = reorder(Player, total_ptspg_diff), x = total_ptspg_diff)) +
  geom_col(fill = "purple") +
  geom_label(aes(label = total_ptspg_diff), size = 5) +
  theme_bw() +
  labs(
    x = "Total Points Per Game Difference",
    y = "Players",
    caption = "Data courtesy of Hockey-Reference"
  )


long_misc_df_exp |> 
  group_by(Player) |> 
  arrange(desc(total_ops_diff)) |> 
  head(n = 10) |> 
  ggplot(aes(y = reorder(Player, total_ops_diff), x = total_ops_diff)) +
  geom_col(fill = "purple") +
  geom_label(aes(label = total_ops_diff), size = 5) +
  theme_bw() +
  labs(
    x = "Total Offensive Point Shares Difference",
    y = "Players",
    caption = "Data courtesy of Hockey-Reference"
  )



long_misc_df_exp |> 
  filter(Pos == "F", num_seasons >= 4, consistency_score != Inf) |> 
  group_by(Player) |> 
  arrange(desc(consistency_score)) |> 
  head(n = 10) |> 
  ggplot(aes(y = reorder(Player, consistency_score), x = consistency_score)) +
  geom_col(fill = "purple") +
  geom_label(aes(label = consistency_score), size = 5) +
  theme_bw() +
  labs(
    x = "Preliminary Consistency Score",
    y = "Players",
    caption = "Data courtesy of Hockey-Reference"
  ) +
  scale_x_continuous(n.breaks = 6)


long_misc_df_exp |> 
  filter(Pos == "F", num_seasons >= 6) |> 
  filter(consistency_score >= 0.5) |> 
  group_by(Player) |> 
  arrange(consistency_score) |> 
  head(n = 10) |> 
  ggplot(aes(y = reorder(Player, -consistency_score), x = consistency_score)) +
  geom_col(fill = "purple") +
  geom_label(aes(label = consistency_score), size = 5) +
  theme_bw() +
  labs(
    x = "Preliminary Consistency Score",
    y = "Players",
    caption = "Data courtesy of Hockey-Reference"
  )
  # scale_x_continuous(limits = c(0, 0.6), n.breaks = 7)

```









