---
title: "Cole's Code"
output: html_document
date: "2024-06-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(broom)
# library(hockeyR)
```



## Thoughts on measuring consistency


- Measure the difference of playing statistics season-over-season
- Weight the metrics differently based on importance
  - consistency score will have more consistent players near 0 & inconsistent players much greater than 0
- playing metrics will have to be per game or per 60 measures for unbiased information


- Use pivot_wider function to increase number of columns in order to take the differences between playing metrics

- Use the average of differences to measure consistency



```{r}
basic_2023_24 <- read_csv("../Data/Basic Stats/basic_stats_23_24.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2023)

basic_2022_23 <- read_csv("../Data/Basic Stats/basic_stats_22_23.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2022)

basic_2021_22 <- read_csv("../Data/Basic Stats/basic_stats_21_22.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2021)

basic_2020_21 <- read_csv("../Data/Basic Stats/basic_stats_20_21.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2020)

basic_2019_20 <- read_csv("../Data/Basic Stats/basic_stats_19_20.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2019)

basic_2018_19 <- read_csv("../Data/Basic Stats/basic_stats_18_19.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2018)

basic_2017_18 <- read_csv("../Data/Basic Stats/basic_stats_17_18.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2017)

basic_2016_17 <- read_csv("../Data/Basic Stats/basic_stats_16_17.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2016)

basic_df <- rbind(basic_2016_17, basic_2017_18, basic_2018_19, basic_2019_20, basic_2020_21, basic_2021_22, basic_2022_23, basic_2023_24)

basic_df <- basic_df |> 
  filter(Pos != "G", GP >= 20)
```



```{r}
long_basic_df <- basic_df |> 
  pivot_wider(id_cols = c(Player, PlayerID), values_from = c(G, A, PTS, EVG, PPG, SHG, EVA, PPA, SHA, `+/-`, TOI, `S%`, PS), names_from = start_year) |> 
  unique()

long_basic_df <- long_basic_df %>% 
  unnest() %>% 
  replace(. == "NULL", NA)

# long_basic_df_exp <- long_basic_df |> 
#   mutate(Gdiff_23_22 = abs(G_2023 - G_2022),
#          Gdiff_22_21 = abs(G_2022 - G_2021),
#          Gdiff_21_20 = abs(G_2021 - G_2020),
#          Gdiff_20_19 = abs(G_2020 - G_2019),
#          Gdiff_19_18 = abs(G_2019 - G_2018),
#          Gdiff_18_17 = abs(G_2018 - G_2017),
#          Gdiff_17_16 = abs(G_2017 - G_2016)
#          ) |>
#   group_by(Player, PlayerID) |>
#   summarise(total_Gdiff = sum(Gdiff_23_22, Gdiff_22_21, Gdiff_21_20, Gdiff_20_19, Gdiff_19_18, Gdiff_18_17, Gdiff_17_16))

long_basic_df <- long_basic_df |> arrange(Player, PlayerID)
```




```{r}
advanced_2023_24 <- read_csv("../Data/Advanced Stats/advanced_stats_23_24.csv") |> 
  mutate(start_year = 2023)

advanced_2022_23 <- read_csv("../Data/Advanced Stats/advanced_stats_22_23.csv") |> 
  mutate(start_year = 2022)

advanced_2021_22 <- read_csv("../Data/Advanced Stats/advanced_stats_21_22.csv") |> 
  mutate(start_year = 2021)

advanced_2020_21 <- read_csv("../Data/Advanced Stats/advanced_stats_20_21.csv") |> 
  mutate(start_year = 2020)

advanced_2019_20 <- read_csv("../Data/Advanced Stats/advanced_stats_19_20.csv") |> 
  mutate(start_year = 2019)

advanced_2018_19 <- read_csv("../Data/Advanced Stats/advanced_stats_18_19.csv") |> 
  mutate(start_year = 2018)

advanced_2017_18 <- read_csv("../Data/Advanced Stats/advanced_stats_17_18.csv") |>
  mutate(start_year = 2017)

advanced_2016_17 <- read_csv("../Data/Advanced Stats/advanced_stats_16_17.csv") |>
  mutate(start_year = 2016)

advanced_df <- rbind(advanced_2016_17, advanced_2017_18, advanced_2018_19, advanced_2019_20, advanced_2020_21, advanced_2021_22, advanced_2022_23, advanced_2023_24)

advanced_df <- advanced_df |> filter(GP >= 20, Pos != "G") |> unique()
```




```{r}
long_adv_df <- advanced_df |> 
  pivot_wider(id_cols = c(Player, PlayerID), values_from = c(GP, CF, CA, FF, FA, `oiSH%`, `oiSV%`, PDO, TK, GV, `CF%`, `FF%`), names_from = start_year) |> 
  unique()

long_adv_df <- long_adv_df %>% 
  unnest() %>% 
  replace(. == "NULL", NA)
```



```{r}
toi_2023_24 <- read_csv("../Data/TOI Stats/toi_stats_23_24.csv") |> 
  mutate(start_year = 2023)

toi_2022_23 <- read_csv("../Data/TOI Stats/toi_stats_22_23.csv") |>
  mutate(start_year = 2022)

toi_2021_22 <- read_csv("../Data/TOI Stats/toi_stats_21_22.csv") |> 
  mutate(start_year = 2021)

toi_2020_21 <- read_csv("../Data/TOI Stats/toi_stats_20_21.csv") |> 
  mutate(start_year = 2020)

toi_2019_20 <- read_csv("../Data/TOI Stats/toi_stats_19_20.csv") |> 
  mutate(start_year = 2019)

toi_2018_19 <- read_csv("../Data/TOI Stats/toi_stats_18_19.csv") |> 
  mutate(start_year = 2018)

toi_2017_18 <- read_csv("../Data/TOI Stats/toi_stats_17_18.csv") |> 
  mutate(start_year = 2017)

toi_2016_17 <- read_csv("../Data/TOI Stats/toi_stats_16_17.csv") |> 
  mutate(start_year = 2016)

toi_df <- rbind(toi_2016_17, toi_2017_18, toi_2018_19, toi_2019_20, toi_2020_21, toi_2021_22, toi_2022_23, toi_2023_24)

toi_df <- toi_df |> select(-`...7`, -`...12`, -`...17`)

toi_df <- toi_df |> filter(Pos != "G", GP >= 20, PlayerID != "ahose02")
```



```{r}
long_toi_df <- toi_df |> 
  pivot_wider(id_cols = c(Player, PlayerID), values_from = c(`TOI (EV)`, `CF% Rel (EV)`, `GF/60 (EV)`, `GA/60 (EV)`, `TOI (PP)`, `CF% Rel (PP)`, `GF/60 (PP)`, `GA/60 (PP)`, `TOI (SH)`, `CF% Rel (SH)`, `GF/60 (SH)`, `GA/60 (SH)`), names_from = start_year) |> 
  unique()

long_toi_df <- long_toi_df %>% 
  unnest() %>%
  replace(. == "NULL", NA)

long_toi_df <- long_toi_df |> arrange(Player, PlayerID)
```



```{r}
misc_2023_24 <- read_csv("../Data/Misc Stats/misc_stats_23_24.csv") |> 
  mutate(start_year = 2023)

misc_2022_23 <- read_csv("../Data/Misc Stats/misc_stats_22_23.csv") |>
  mutate(start_year = 2022)

misc_2021_22 <- read_csv("../Data/Misc Stats/misc_stats_21_22.csv") |> 
  mutate(start_year = 2021)

misc_2020_21 <- read_csv("../Data/Misc Stats/misc_stats_20_21.csv") |> 
  mutate(start_year = 2020)

misc_2019_20 <- read_csv("../Data/Misc Stats/misc_stats_19_20.csv") |> 
  mutate(start_year = 2019)

misc_2018_19 <- read_csv("../Data/Misc Stats/misc_stats_18_19.csv") |> 
  mutate(start_year = 2018)

misc_2017_18 <- read_csv("../Data/Misc Stats/misc_stats_17_18.csv") |> 
  mutate(start_year = 2017)

misc_2016_17 <- read_csv("../Data/Misc Stats/misc_stats_16_17.csv") |> 
  mutate(start_year = 2016)

misc_df <- rbind(misc_2016_17, misc_2017_18, misc_2018_19, misc_2019_20, misc_2020_21, misc_2021_22, misc_2022_23, misc_2023_24)

misc_df <- misc_df |> filter(Pos != "G", GP >= 20)
```



```{r}
misc_df_pos <- misc_df |> 
  mutate(Pos = str_extract(Pos, "[A-Z]"))

num_seasons <- misc_df |> 
  group_by(Player, PlayerID) |> 
  count() |> 
  ungroup() |> 
  mutate(num_seasons = n) |> 
  unique()

player_pos <- misc_df_pos |> 
  group_by(Player, PlayerID) |> 
  summarise(Pos = ifelse(Pos %in% c("C", "L", "R", "W", "F"), "F", "D")) |> 
  unique()
```



```{r}
long_misc_df <- misc_df |> 
  pivot_wider(id_cols = c(Player, PlayerID), values_from = c(`G/GP`, `A/GP`, `PTS/GP`, `GC/GP`, `S/GP`, `+/-`, `OPS`, `DPS`, `PS`), names_from = start_year)

long_misc_df <- long_misc_df %>% 
  unnest() %>% 
  replace(. == "NULL", NA)
  # mutate(across(`G/GP_2023`:`PS_2016`, ~replace_na(.x, 0)))

# long_misc_df_exp <- long_misc_df |>
#   group_by(Player) |> 
#   summarise(ptspg_diff_23_22 = abs(`PTS/GP_2023` - `PTS/GP_2022`),
#          ptspg_diff_22_21 = abs(`PTS/GP_2022` - `PTS/GP_2021`),
#          ptspg_diff_21_20 = abs(`PTS/GP_2021` - `PTS/GP_2020`),
#          ptspg_diff_20_19 = abs(`PTS/GP_2020` - `PTS/GP_2019`),
#          ptspg_diff_19_18 = abs(`PTS/GP_2019` - `PTS/GP_2018`),
#          ptspg_diff_18_17 = abs(`PTS/GP_2018` - `PTS/GP_2017`),
#          ptspg_diff_17_16 = abs(`PTS/GP_2017` - `PTS/GP_2016`),
#          
#          ops_diff_23_22 = abs(OPS_2023 - OPS_2022),
#          ops_diff_22_21 = abs(OPS_2022 - OPS_2021),
#          ops_diff_21_20 = abs(OPS_2021 - OPS_2020),
#          ops_diff_20_19 = abs(OPS_2020 - OPS_2019),
#          ops_diff_19_18 = abs(OPS_2019 - OPS_2018),
#          ops_diff_18_17 = abs(OPS_2018 - OPS_2017),
#          ops_diff_17_16 = abs(OPS_2017 - OPS_2016),
#          
#          p_m_diff_23_22 = abs(`+/-_2023` - `+/-_2022`),
#          p_m_diff_22_21 = abs(`+/-_2022` - `+/-_2021`),
#          p_m_diff_21_20 = abs(`+/-_2021` - `+/-_2020`),
#          p_m_diff_20_19 = abs(`+/-_2020` - `+/-_2019`),
#          p_m_diff_19_18 = abs(`+/-_2019` - `+/-_2018`),
#          p_m_diff_18_17 = abs(`+/-_2018` - `+/-_2017`),
#          p_m_diff_17_16 = abs(`+/-_2017` - `+/-_2016`),
#          
#          total_ptspg_diff = sum(ptspg_diff_23_22, ptspg_diff_22_21, ptspg_diff_21_20, ptspg_diff_20_19, ptspg_diff_19_18, ptspg_diff_18_17, ptspg_diff_17_16, na.rm = TRUE),
#          
#          total_ops_diff = sum(ops_diff_23_22, ops_diff_22_21, ops_diff_21_20, ops_diff_20_19, ops_diff_19_18, ops_diff_18_17, ops_diff_17_16, na.rm = TRUE),
#          
#          total_p_m_diff = sum(p_m_diff_23_22, p_m_diff_22_21, p_m_diff_21_20, p_m_diff_20_19, p_m_diff_19_18, p_m_diff_18_17, p_m_diff_17_16, na.rm = TRUE)
#   )
```



```{r, Making Complete Dataset}
complete_df3 <- cbind(long_basic_df, long_adv_df, long_toi_df, long_misc_df)

complete_df <- complete_df3[ , !duplicated(colnames(complete_df3))]

complete_df <- complete_df |> full_join(player_pos)

complete_df <- complete_df |> full_join(num_seasons) |> select(-n)


complete_df_sel <- complete_df |> 
  filter(num_seasons >= 5)
```



```{r}
playmaker <- complete_df_sel |> select(Player, PlayerID, starts_with("TOI_"), contains("EVA_"), contains("PPA_"), contains("SHA_"))

playmaker2 <- playmaker |> 
  group_by(Player, PlayerID) |> 
  summarise(
    `EVA/60_2023` = (EVA_2023 / TOI_2023) * 60,
    `EVA/60_2022` = (EVA_2022 / TOI_2022) * 60,
    `EVA/60_2021` = (EVA_2021 / TOI_2021) * 60,
    `EVA/60_2020` = (EVA_2020 / TOI_2020) * 60,
    `EVA/60_2019` = (EVA_2019 / TOI_2019) * 60,
    `EVA/60_2018` = (EVA_2018 / TOI_2018) * 60,
    `EVA/60_2017` = (EVA_2017 / TOI_2017) * 60,
    `EVA/60_2016` = (EVA_2016 / TOI_2016) * 60,
    
    `PPA/60_2023` = (PPA_2023 / TOI_2023) * 60,
    `PPA/60_2022` = (PPA_2022 / TOI_2022) * 60,
    `PPA/60_2021` = (PPA_2021 / TOI_2021) * 60,
    `PPA/60_2020` = (PPA_2020 / TOI_2020) * 60,
    `PPA/60_2019` = (PPA_2019 / TOI_2019) * 60,
    `PPA/60_2018` = (PPA_2018 / TOI_2018) * 60,
    `PPA/60_2017` = (PPA_2017 / TOI_2017) * 60,
    `PPA/60_2016` = (PPA_2016 / TOI_2016) * 60,
    
    `SHA/60_2023` = (SHA_2023 / TOI_2023) * 60,
    `SHA/60_2022` = (SHA_2022 / TOI_2022) * 60,
    `SHA/60_2021` = (SHA_2021 / TOI_2021) * 60,
    `SHA/60_2020` = (SHA_2020 / TOI_2020) * 60,
    `SHA/60_2019` = (SHA_2019 / TOI_2019) * 60,
    `SHA/60_2018` = (SHA_2018 / TOI_2018) * 60,
    `SHA/60_2017` = (SHA_2017 / TOI_2017) * 60,
    `SHA/60_2016` = (SHA_2016 / TOI_2016) * 60,
         )

playmaker3 <- playmaker2 |> 
  ungroup() |> 
  mutate(mean_EVA_23 = mean(`EVA/60_2023`, na.rm = TRUE),
         mean_EVA_22 = mean(`EVA/60_2022`, na.rm = TRUE),
         mean_EVA_21 = mean(`EVA/60_2021`, na.rm = TRUE),
         mean_EVA_20 = mean(`EVA/60_2020`, na.rm = TRUE),
         mean_EVA_19 = mean(`EVA/60_2019`, na.rm = TRUE),
         mean_EVA_18 = mean(`EVA/60_2018`, na.rm = TRUE),
         mean_EVA_17 = mean(`EVA/60_2017`, na.rm = TRUE),
         mean_EVA_16 = mean(`EVA/60_2016`, na.rm = TRUE),
         
         mean_PPA_23 = mean(`PPA/60_2023`, na.rm = TRUE),
         mean_PPA_22 = mean(`PPA/60_2022`, na.rm = TRUE),
         mean_PPA_21 = mean(`PPA/60_2021`, na.rm = TRUE),
         mean_PPA_20 = mean(`PPA/60_2020`, na.rm = TRUE),
         mean_PPA_19 = mean(`PPA/60_2019`, na.rm = TRUE),
         mean_PPA_18 = mean(`PPA/60_2018`, na.rm = TRUE),
         mean_PPA_17 = mean(`PPA/60_2017`, na.rm = TRUE),
         mean_PPA_16 = mean(`PPA/60_2016`, na.rm = TRUE),
         
         mean_SHA_23 = mean(`SHA/60_2023`, na.rm = TRUE),
         mean_SHA_22 = mean(`SHA/60_2022`, na.rm = TRUE),
         mean_SHA_21 = mean(`SHA/60_2021`, na.rm = TRUE),
         mean_SHA_20 = mean(`SHA/60_2020`, na.rm = TRUE),
         mean_SHA_19 = mean(`SHA/60_2019`, na.rm = TRUE),
         mean_SHA_18 = mean(`SHA/60_2018`, na.rm = TRUE),
         mean_SHA_17 = mean(`SHA/60_2017`, na.rm = TRUE),
         mean_SHA_16 = mean(`SHA/60_2016`, na.rm = TRUE),
         
         
         sd_EVA_23 = sd(`EVA/60_2023`, na.rm = TRUE),
         sd_EVA_22 = sd(`EVA/60_2022`, na.rm = TRUE),
         sd_EVA_21 = sd(`EVA/60_2021`, na.rm = TRUE),
         sd_EVA_20 = sd(`EVA/60_2020`, na.rm = TRUE),
         sd_EVA_19 = sd(`EVA/60_2019`, na.rm = TRUE),
         sd_EVA_18 = sd(`EVA/60_2018`, na.rm = TRUE),
         sd_EVA_17 = sd(`EVA/60_2017`, na.rm = TRUE),
         sd_EVA_16 = sd(`EVA/60_2016`, na.rm = TRUE),
         
         sd_PPA_23 = sd(`PPA/60_2023`, na.rm = TRUE),
         sd_PPA_22 = sd(`PPA/60_2022`, na.rm = TRUE),
         sd_PPA_21 = sd(`PPA/60_2021`, na.rm = TRUE),
         sd_PPA_20 = sd(`PPA/60_2020`, na.rm = TRUE),
         sd_PPA_19 = sd(`PPA/60_2019`, na.rm = TRUE),
         sd_PPA_18 = sd(`PPA/60_2018`, na.rm = TRUE),
         sd_PPA_17 = sd(`PPA/60_2017`, na.rm = TRUE),
         sd_PPA_16 = sd(`PPA/60_2016`, na.rm = TRUE),
         
         sd_SHA_23 = sd(`SHA/60_2023`, na.rm = TRUE),
         sd_SHA_22 = sd(`SHA/60_2022`, na.rm = TRUE),
         sd_SHA_21 = sd(`SHA/60_2021`, na.rm = TRUE),
         sd_SHA_20 = sd(`SHA/60_2020`, na.rm = TRUE),
         sd_SHA_19 = sd(`SHA/60_2019`, na.rm = TRUE),
         sd_SHA_18 = sd(`SHA/60_2018`, na.rm = TRUE),
         sd_SHA_17 = sd(`SHA/60_2017`, na.rm = TRUE),
         sd_SHA_16 = sd(`SHA/60_2016`, na.rm = TRUE),
         )


playmaker_score <- playmaker3 |> 
  group_by(Player, PlayerID) |> 
  summarise(
    EVA_zscore_2023 = (`EVA/60_2023` - `mean_EVA_23`) / `sd_EVA_23`,
    EVA_zscore_2022 = (`EVA/60_2022` - `mean_EVA_22`) / `sd_EVA_22`,
    EVA_zscore_2021 = (`EVA/60_2021` - `mean_EVA_21`) / `sd_EVA_21`,
    EVA_zscore_2020 = (`EVA/60_2020` - `mean_EVA_20`) / `sd_EVA_20`,
    EVA_zscore_2019 = (`EVA/60_2019` - `mean_EVA_19`) / `sd_EVA_19`,
    EVA_zscore_2018 = (`EVA/60_2018` - `mean_EVA_18`) / `sd_EVA_18`,
    EVA_zscore_2017 = (`EVA/60_2017` - `mean_EVA_17`) / `sd_EVA_17`,
    EVA_zscore_2016 = (`EVA/60_2016` - `mean_EVA_16`) / `sd_EVA_16`,
    
    PPA_zscore_2023 = (`PPA/60_2023` - `mean_PPA_23`) / `sd_PPA_23`,
    PPA_zscore_2022 = (`PPA/60_2022` - `mean_PPA_22`) / `sd_PPA_22`,
    PPA_zscore_2021 = (`PPA/60_2021` - `mean_PPA_21`) / `sd_PPA_21`,
    PPA_zscore_2020 = (`PPA/60_2020` - `mean_PPA_20`) / `sd_PPA_20`,
    PPA_zscore_2019 = (`PPA/60_2019` - `mean_PPA_19`) / `sd_PPA_19`,
    PPA_zscore_2018 = (`PPA/60_2018` - `mean_PPA_18`) / `sd_PPA_18`,
    PPA_zscore_2017 = (`PPA/60_2017` - `mean_PPA_17`) / `sd_PPA_17`,
    PPA_zscore_2016 = (`PPA/60_2016` - `mean_PPA_16`) / `sd_PPA_16`,
    
    SHA_zscore_2023 = (`SHA/60_2023` - `mean_SHA_23`) / `sd_SHA_23`,
    SHA_zscore_2022 = (`SHA/60_2022` - `mean_SHA_22`) / `sd_SHA_22`,
    SHA_zscore_2021 = (`SHA/60_2021` - `mean_SHA_21`) / `sd_SHA_21`,
    SHA_zscore_2020 = (`SHA/60_2020` - `mean_SHA_20`) / `sd_SHA_20`,
    SHA_zscore_2019 = (`SHA/60_2019` - `mean_SHA_19`) / `sd_SHA_19`,
    SHA_zscore_2018 = (`SHA/60_2018` - `mean_SHA_18`) / `sd_SHA_18`,
    SHA_zscore_2017 = (`SHA/60_2017` - `mean_SHA_17`) / `sd_SHA_17`,
    SHA_zscore_2016 = (`SHA/60_2016` - `mean_SHA_16`) / `sd_SHA_16`
            )


```



```{r}
master_playmaker <- playmaker_score |> 
  select(-PlayerID) |> 
  pivot_longer(-c(Player),
               names_to = c(".value", "Year"), 
               names_sep = "_zscore_")
```



```{r}
master_playmaker |> 
  mutate(Groupings = ifelse(EVA <= 0.5 & EVA >= -0.5, "Consistent", "Inconsistent")) |> 
  ggplot(aes(x = Year, y = EVA)) +
  geom_point(aes(fill = Groupings)) +
  geom_hline(yintercept = 0.5, fill = "red", linetype = "dashed")
```





```{r}
long_misc_df_exp |> 
  group_by(Player) |> 
  arrange(desc(total_ptspg_diff)) |> 
  head(n = 10) |> 
  ggplot(aes(y = reorder(Player, total_ptspg_diff), x = total_ptspg_diff)) +
  geom_col(fill = "purple") +
  geom_label(aes(label = total_ptspg_diff), size = 5) +
  theme_bw() +
  labs(
    x = "Total Points Per Game Difference",
    y = "Players",
    caption = "Data courtesy of Hockey-Reference"
  )


long_misc_df_exp |> 
  group_by(Player) |> 
  arrange(desc(total_ops_diff)) |> 
  head(n = 10) |> 
  ggplot(aes(y = reorder(Player, total_ops_diff), x = total_ops_diff)) +
  geom_col(fill = "purple") +
  geom_label(aes(label = total_ops_diff), size = 5) +
  theme_bw() +
  labs(
    x = "Total Offensive Point Shares Difference",
    y = "Players",
    caption = "Data courtesy of Hockey-Reference"
  )



# long_misc_df_exp |> 
#   filter(Pos == "F", num_seasons >= 4, consistency_score != Inf) |> 
#   group_by(Player) |> 
#   arrange(desc(consistency_score)) |> 
#   head(n = 10) |> 
#   ggplot(aes(y = reorder(Player, consistency_score), x = consistency_score)) +
#   geom_col(fill = "purple") +
#   geom_label(aes(label = consistency_score), size = 5) +
#   theme_bw() +
#   labs(
#     x = "Preliminary Consistency Score",
#     y = "Players",
#     caption = "Data courtesy of Hockey-Reference"
#   ) +
#   scale_x_continuous(n.breaks = 6)
# 
# 
# long_misc_df_exp |> 
#   filter(Pos == "F", num_seasons >= 6) |> 
#   filter(consistency_score >= 0.5) |> 
#   group_by(Player) |> 
#   arrange(consistency_score) |> 
#   head(n = 10) |> 
#   ggplot(aes(y = reorder(Player, -consistency_score), x = consistency_score)) +
#   geom_col(fill = "purple") +
#   geom_label(aes(label = consistency_score), size = 5) +
#   theme_bw() +
#   labs(
#     x = "Preliminary Consistency Score",
#     y = "Players",
#     caption = "Data courtesy of Hockey-Reference"
#   )
  # scale_x_continuous(limits = c(0, 0.6), n.breaks = 7)

```




























