---
title: "Cole's Code"
output: html_document
date: "2024-06-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(broom)
# library(hockeyR)
```



## Thoughts on measuring consistency


- Measure the difference of playing statistics season-over-season
- Weight the metrics differently based on importance
  - consistency score will have more consistent players near 0 & inconsistent players much greater than 0
- playing metrics will have to be per game or per 60 measures for unbiased information


- Use pivot_wider function to increase number of columns in order to take the differences between playing metrics

- Use the average of differences to measure consistency



```{r}
basic_2023_24 <- read_csv("../Data/Basic Stats/basic_stats_23_24.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2023)

basic_2022_23 <- read_csv("../Data/Basic Stats/basic_stats_22_23.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2022)

basic_2021_22 <- read_csv("../Data/Basic Stats/basic_stats_21_22.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2021)

basic_2020_21 <- read_csv("../Data/Basic Stats/basic_stats_20_21.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2020)

basic_2019_20 <- read_csv("../Data/Basic Stats/basic_stats_19_20.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2019)

basic_2018_19 <- read_csv("../Data/Basic Stats/basic_stats_18_19.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2018)

basic_2017_18 <- read_csv("../Data/Basic Stats/basic_stats_17_18.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2017)

basic_2016_17 <- read_csv("../Data/Basic Stats/basic_stats_16_17.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |> mutate(start_year = 2016)

basic_df <- rbind(basic_2016_17, basic_2017_18, basic_2018_19, basic_2019_20, basic_2020_21, basic_2021_22, basic_2022_23, basic_2023_24)

basic_df <- basic_df |> 
  filter(Pos != "G", GP >= 20)
```



```{r}
long_basic_df <- basic_df |> 
  pivot_wider(id_cols = c(Player, PlayerID), values_from = c(G, A, PTS, EVG, PPG, SHG, EVA, PPA, SHA, `+/-`, TOI, `S%`, PS, HIT, BLK, PIM), names_from = start_year) |> 
  unique()

long_basic_df <- long_basic_df %>% 
  unnest() %>% 
  replace(. == "NULL", NA)

# long_basic_df_exp <- long_basic_df |> 
#   mutate(Gdiff_23_22 = abs(G_2023 - G_2022),
#          Gdiff_22_21 = abs(G_2022 - G_2021),
#          Gdiff_21_20 = abs(G_2021 - G_2020),
#          Gdiff_20_19 = abs(G_2020 - G_2019),
#          Gdiff_19_18 = abs(G_2019 - G_2018),
#          Gdiff_18_17 = abs(G_2018 - G_2017),
#          Gdiff_17_16 = abs(G_2017 - G_2016)
#          ) |>
#   group_by(Player, PlayerID) |>
#   summarise(total_Gdiff = sum(Gdiff_23_22, Gdiff_22_21, Gdiff_21_20, Gdiff_20_19, Gdiff_19_18, Gdiff_18_17, Gdiff_17_16))

long_basic_df <- long_basic_df |> arrange(Player, PlayerID)
```




```{r}
advanced_2023_24 <- read_csv("../Data/Advanced Stats/advanced_stats_23_24.csv") |> 
  mutate(start_year = 2023)

advanced_2022_23 <- read_csv("../Data/Advanced Stats/advanced_stats_22_23.csv") |> 
  mutate(start_year = 2022)

advanced_2021_22 <- read_csv("../Data/Advanced Stats/advanced_stats_21_22.csv") |> 
  mutate(start_year = 2021)

advanced_2020_21 <- read_csv("../Data/Advanced Stats/advanced_stats_20_21.csv") |> 
  mutate(start_year = 2020)

advanced_2019_20 <- read_csv("../Data/Advanced Stats/advanced_stats_19_20.csv") |> 
  mutate(start_year = 2019)

advanced_2018_19 <- read_csv("../Data/Advanced Stats/advanced_stats_18_19.csv") |> 
  mutate(start_year = 2018)

advanced_2017_18 <- read_csv("../Data/Advanced Stats/advanced_stats_17_18.csv") |>
  mutate(start_year = 2017)

advanced_2016_17 <- read_csv("../Data/Advanced Stats/advanced_stats_16_17.csv") |>
  mutate(start_year = 2016)

advanced_df <- rbind(advanced_2016_17, advanced_2017_18, advanced_2018_19, advanced_2019_20, advanced_2020_21, advanced_2021_22, advanced_2022_23, advanced_2023_24)

advanced_df <- advanced_df |> filter(GP >= 20, Pos != "G") |> unique()
```




```{r}
long_adv_df <- advanced_df |> 
  pivot_wider(id_cols = c(Player, PlayerID), values_from = c(GP, CF, CA, FF, FA, `oiSH%`, `oiSV%`, PDO, TK, GV, `CF%`, `FF%`), names_from = start_year) |> 
  unique()

long_adv_df <- long_adv_df %>% 
  unnest() %>% 
  replace(. == "NULL", NA)

long_adv_df <- long_adv_df |> arrange(Player, PlayerID)
```



```{r}
toi_2023_24 <- read_csv("../Data/TOI Stats/toi_stats_23_24.csv") |> 
  mutate(start_year = 2023)

toi_2022_23 <- read_csv("../Data/TOI Stats/toi_stats_22_23.csv") |>
  mutate(start_year = 2022)

toi_2021_22 <- read_csv("../Data/TOI Stats/toi_stats_21_22.csv") |> 
  mutate(start_year = 2021)

toi_2020_21 <- read_csv("../Data/TOI Stats/toi_stats_20_21.csv") |> 
  mutate(start_year = 2020)

toi_2019_20 <- read_csv("../Data/TOI Stats/toi_stats_19_20.csv") |> 
  mutate(start_year = 2019)

toi_2018_19 <- read_csv("../Data/TOI Stats/toi_stats_18_19.csv") |> 
  mutate(start_year = 2018)

toi_2017_18 <- read_csv("../Data/TOI Stats/toi_stats_17_18.csv") |> 
  mutate(start_year = 2017)

toi_2016_17 <- read_csv("../Data/TOI Stats/toi_stats_16_17.csv") |> 
  mutate(start_year = 2016)

toi_df <- rbind(toi_2016_17, toi_2017_18, toi_2018_19, toi_2019_20, toi_2020_21, toi_2021_22, toi_2022_23, toi_2023_24)

toi_df <- toi_df |> select(-`...7`, -`...12`, -`...17`)

toi_df <- toi_df |> filter(Pos != "G", GP >= 20, PlayerID != "ahose02")
```



```{r}
long_toi_df <- toi_df |> 
  pivot_wider(id_cols = c(Player, PlayerID), values_from = c(`TOI (EV)`, `CF% Rel (EV)`, `GF/60 (EV)`, `GA/60 (EV)`, `TOI (PP)`, `CF% Rel (PP)`, `GF/60 (PP)`, `GA/60 (PP)`, `TOI (SH)`, `CF% Rel (SH)`, `GF/60 (SH)`, `GA/60 (SH)`), names_from = start_year) |> 
  unique()

long_toi_df <- long_toi_df %>% 
  unnest() %>%
  replace(. == "NULL", NA)

long_toi_df <- long_toi_df |> arrange(Player, PlayerID)
```



```{r}
misc_2023_24 <- read_csv("../Data/Misc Stats/misc_stats_23_24.csv") |> 
  mutate(start_year = 2023)

misc_2022_23 <- read_csv("../Data/Misc Stats/misc_stats_22_23.csv") |>
  mutate(start_year = 2022)

misc_2021_22 <- read_csv("../Data/Misc Stats/misc_stats_21_22.csv") |> 
  mutate(start_year = 2021)

misc_2020_21 <- read_csv("../Data/Misc Stats/misc_stats_20_21.csv") |> 
  mutate(start_year = 2020)

misc_2019_20 <- read_csv("../Data/Misc Stats/misc_stats_19_20.csv") |> 
  mutate(start_year = 2019)

misc_2018_19 <- read_csv("../Data/Misc Stats/misc_stats_18_19.csv") |> 
  mutate(start_year = 2018)

misc_2017_18 <- read_csv("../Data/Misc Stats/misc_stats_17_18.csv") |> 
  mutate(start_year = 2017)

misc_2016_17 <- read_csv("../Data/Misc Stats/misc_stats_16_17.csv") |> 
  mutate(start_year = 2016)

misc_df <- rbind(misc_2016_17, misc_2017_18, misc_2018_19, misc_2019_20, misc_2020_21, misc_2021_22, misc_2022_23, misc_2023_24)

misc_df <- misc_df |> filter(Pos != "G", GP >= 20)
```



```{r}
misc_df_pos <- misc_df |> 
  mutate(Pos = str_extract(Pos, "[A-Z]"))

num_seasons <- misc_df |> 
  group_by(Player, PlayerID) |> 
  count() |> 
  ungroup() |> 
  mutate(num_seasons = n) |> 
  unique()

player_pos <- misc_df_pos |> 
  group_by(Player, PlayerID) |> 
  summarise(Pos = ifelse(Pos %in% c("C", "L", "R", "W", "F"), "F", "D")) |> 
  unique()
```



```{r}
long_misc_df <- misc_df |> 
  pivot_wider(id_cols = c(Player, PlayerID), values_from = c(GC, `G/GP`, `A/GP`, `PTS/GP`, `GC/GP`, `S/GP`, `+/-`, `OPS`, `DPS`, `PS`), names_from = start_year)

long_misc_df <- long_misc_df %>% 
  unnest() %>% 
  replace(. == "NULL", NA)
  # mutate(across(`G/GP_2023`:`PS_2016`, ~replace_na(.x, 0)))

long_misc_df <- long_misc_df |> arrange(Player, PlayerID)

# long_misc_df_exp <- long_misc_df |>
#   group_by(Player) |> 
#   summarise(ptspg_diff_23_22 = abs(`PTS/GP_2023` - `PTS/GP_2022`),
#          ptspg_diff_22_21 = abs(`PTS/GP_2022` - `PTS/GP_2021`),
#          ptspg_diff_21_20 = abs(`PTS/GP_2021` - `PTS/GP_2020`),
#          ptspg_diff_20_19 = abs(`PTS/GP_2020` - `PTS/GP_2019`),
#          ptspg_diff_19_18 = abs(`PTS/GP_2019` - `PTS/GP_2018`),
#          ptspg_diff_18_17 = abs(`PTS/GP_2018` - `PTS/GP_2017`),
#          ptspg_diff_17_16 = abs(`PTS/GP_2017` - `PTS/GP_2016`),
#          
#          ops_diff_23_22 = abs(OPS_2023 - OPS_2022),
#          ops_diff_22_21 = abs(OPS_2022 - OPS_2021),
#          ops_diff_21_20 = abs(OPS_2021 - OPS_2020),
#          ops_diff_20_19 = abs(OPS_2020 - OPS_2019),
#          ops_diff_19_18 = abs(OPS_2019 - OPS_2018),
#          ops_diff_18_17 = abs(OPS_2018 - OPS_2017),
#          ops_diff_17_16 = abs(OPS_2017 - OPS_2016),
#          
#          p_m_diff_23_22 = abs(`+/-_2023` - `+/-_2022`),
#          p_m_diff_22_21 = abs(`+/-_2022` - `+/-_2021`),
#          p_m_diff_21_20 = abs(`+/-_2021` - `+/-_2020`),
#          p_m_diff_20_19 = abs(`+/-_2020` - `+/-_2019`),
#          p_m_diff_19_18 = abs(`+/-_2019` - `+/-_2018`),
#          p_m_diff_18_17 = abs(`+/-_2018` - `+/-_2017`),
#          p_m_diff_17_16 = abs(`+/-_2017` - `+/-_2016`),
#          
#          total_ptspg_diff = sum(ptspg_diff_23_22, ptspg_diff_22_21, ptspg_diff_21_20, ptspg_diff_20_19, ptspg_diff_19_18, ptspg_diff_18_17, ptspg_diff_17_16, na.rm = TRUE),
#          
#          total_ops_diff = sum(ops_diff_23_22, ops_diff_22_21, ops_diff_21_20, ops_diff_20_19, ops_diff_19_18, ops_diff_18_17, ops_diff_17_16, na.rm = TRUE),
#          
#          total_p_m_diff = sum(p_m_diff_23_22, p_m_diff_22_21, p_m_diff_21_20, p_m_diff_20_19, p_m_diff_19_18, p_m_diff_18_17, p_m_diff_17_16, na.rm = TRUE)
#   )
```



```{r}
complete_df1 <- cbind(basic_df, advanced_df, toi_df, misc_df)

complete_df <- complete_df1[ , !duplicated(colnames(complete_df1))]

complete_df <- complete_df |> filter(Player != "Daniel Sedin", Player != "Henrik Sedin", Player != "Jarome Iginla", Player != "Marián Hossa")

complete_df <- complete_df |> full_join(player_pos)

complete_df <- complete_df |> full_join(num_seasons) |> select(-n)

complete_df <- complete_df |> 
  filter(num_seasons >= 5)
```





```{r}
feat <- complete_df |> select(HIT, BLK, PIM) |> na.omit()

# glimpse(feat)

nhl_pca <- prcomp(feat, center = TRUE, scale. = TRUE)

# summary(nhl_pca)

library(factoextra)
# fviz_pca_var(): projection of variables
# fviz_pca_ind(): display observations with first two PCs
nhl_pca |> 
  fviz_pca_biplot(label = "var",
                  alpha.ind = 0.25,
                  alpha.var = 0.75,
                  col.var = "darkblue",
                  repel = TRUE)

nhl_pca |> 
  fviz_eig(addlabels = TRUE) +
  geom_hline(yintercept = 100 * (1 / ncol(nhl_pca$x)), linetype = "dashed", color = "darkred")

library(broom)
nhl_pca |>
  tidy(matrix = "eigenvalues") |>
  ggplot(aes(x = PC, y = cumulative)) +
  geom_line() + 
  geom_point() +
  scale_x_continuous(breaks = 1:ncol(nhl_pca$x))
```






```{r}
complete_df_long1 <- cbind(long_basic_df, long_adv_df, long_toi_df, long_misc_df)

complete_df_long <- complete_df_long1[ , !duplicated(colnames(complete_df_long1))]

complete_df_long <- complete_df_long |> filter(Player != "Daniel Sedin", Player != "Henrik Sedin", Player != "Jarome Iginla", Player != "Marián Hossa")

complete_df_long <- complete_df_long |> full_join(player_pos)

complete_df_long <- complete_df_long |> full_join(num_seasons) |> select(-n)

complete_df_long <- complete_df_long |> 
  filter(num_seasons >= 5)
```



## Playmaker Consistency Metric



```{r}
playmaker <- complete_df_long |> select(Player, PlayerID, starts_with("GC"), starts_with("TOI_"), contains("EVA_"), contains("PPA_"))

playmaker2 <- playmaker |> 
  group_by(Player, PlayerID) |> 
  summarise(
    `EVA/60_2023` = (EVA_2023 / TOI_2023) * 60,
    `EVA/60_2022` = (EVA_2022 / TOI_2022) * 60,
    `EVA/60_2021` = (EVA_2021 / TOI_2021) * 60,
    `EVA/60_2020` = (EVA_2020 / TOI_2020) * 60,
    `EVA/60_2019` = (EVA_2019 / TOI_2019) * 60,
    `EVA/60_2018` = (EVA_2018 / TOI_2018) * 60,
    `EVA/60_2017` = (EVA_2017 / TOI_2017) * 60,
    `EVA/60_2016` = (EVA_2016 / TOI_2016) * 60,
    
    `PPA/60_2023` = (PPA_2023 / TOI_2023) * 60,
    `PPA/60_2022` = (PPA_2022 / TOI_2022) * 60,
    `PPA/60_2021` = (PPA_2021 / TOI_2021) * 60,
    `PPA/60_2020` = (PPA_2020 / TOI_2020) * 60,
    `PPA/60_2019` = (PPA_2019 / TOI_2019) * 60,
    `PPA/60_2018` = (PPA_2018 / TOI_2018) * 60,
    `PPA/60_2017` = (PPA_2017 / TOI_2017) * 60,
    `PPA/60_2016` = (PPA_2016 / TOI_2016) * 60,
    
    `GC/60_2023` = (GC_2023 / TOI_2023) * 60,
    `GC/60_2022` = (GC_2022 / TOI_2022) * 60,
    `GC/60_2021` = (GC_2021 / TOI_2021) * 60,
    `GC/60_2020` = (GC_2020 / TOI_2020) * 60,
    `GC/60_2019` = (GC_2019 / TOI_2019) * 60,
    `GC/60_2018` = (GC_2018 / TOI_2018) * 60,
    `GC/60_2017` = (GC_2017 / TOI_2017) * 60,
    `GC/60_2016` = (GC_2016 / TOI_2016) * 60,
         )

playmaker3 <- playmaker2 |> 
  ungroup() |> 
  mutate(mean_EVA_23 = mean(`EVA/60_2023`, na.rm = TRUE),
         mean_EVA_22 = mean(`EVA/60_2022`, na.rm = TRUE),
         mean_EVA_21 = mean(`EVA/60_2021`, na.rm = TRUE),
         mean_EVA_20 = mean(`EVA/60_2020`, na.rm = TRUE),
         mean_EVA_19 = mean(`EVA/60_2019`, na.rm = TRUE),
         mean_EVA_18 = mean(`EVA/60_2018`, na.rm = TRUE),
         mean_EVA_17 = mean(`EVA/60_2017`, na.rm = TRUE),
         mean_EVA_16 = mean(`EVA/60_2016`, na.rm = TRUE),
         
         mean_PPA_23 = mean(`PPA/60_2023`, na.rm = TRUE),
         mean_PPA_22 = mean(`PPA/60_2022`, na.rm = TRUE),
         mean_PPA_21 = mean(`PPA/60_2021`, na.rm = TRUE),
         mean_PPA_20 = mean(`PPA/60_2020`, na.rm = TRUE),
         mean_PPA_19 = mean(`PPA/60_2019`, na.rm = TRUE),
         mean_PPA_18 = mean(`PPA/60_2018`, na.rm = TRUE),
         mean_PPA_17 = mean(`PPA/60_2017`, na.rm = TRUE),
         mean_PPA_16 = mean(`PPA/60_2016`, na.rm = TRUE),
         
         mean_GC_23 = mean(`GC/60_2023`, na.rm = TRUE),
         mean_GC_22 = mean(`GC/60_2022`, na.rm = TRUE),
         mean_GC_21 = mean(`GC/60_2021`, na.rm = TRUE),
         mean_GC_20 = mean(`GC/60_2020`, na.rm = TRUE),
         mean_GC_19 = mean(`GC/60_2019`, na.rm = TRUE),
         mean_GC_18 = mean(`GC/60_2018`, na.rm = TRUE),
         mean_GC_17 = mean(`GC/60_2017`, na.rm = TRUE),
         mean_GC_16 = mean(`GC/60_2016`, na.rm = TRUE),
         
         
         sd_EVA_23 = sd(`EVA/60_2023`, na.rm = TRUE),
         sd_EVA_22 = sd(`EVA/60_2022`, na.rm = TRUE),
         sd_EVA_21 = sd(`EVA/60_2021`, na.rm = TRUE),
         sd_EVA_20 = sd(`EVA/60_2020`, na.rm = TRUE),
         sd_EVA_19 = sd(`EVA/60_2019`, na.rm = TRUE),
         sd_EVA_18 = sd(`EVA/60_2018`, na.rm = TRUE),
         sd_EVA_17 = sd(`EVA/60_2017`, na.rm = TRUE),
         sd_EVA_16 = sd(`EVA/60_2016`, na.rm = TRUE),
         
         sd_PPA_23 = sd(`PPA/60_2023`, na.rm = TRUE),
         sd_PPA_22 = sd(`PPA/60_2022`, na.rm = TRUE),
         sd_PPA_21 = sd(`PPA/60_2021`, na.rm = TRUE),
         sd_PPA_20 = sd(`PPA/60_2020`, na.rm = TRUE),
         sd_PPA_19 = sd(`PPA/60_2019`, na.rm = TRUE),
         sd_PPA_18 = sd(`PPA/60_2018`, na.rm = TRUE),
         sd_PPA_17 = sd(`PPA/60_2017`, na.rm = TRUE),
         sd_PPA_16 = sd(`PPA/60_2016`, na.rm = TRUE),
         
         sd_GC_23 = sd(`GC/60_2023`, na.rm = TRUE),
         sd_GC_22 = sd(`GC/60_2022`, na.rm = TRUE),
         sd_GC_21 = sd(`GC/60_2021`, na.rm = TRUE),
         sd_GC_20 = sd(`GC/60_2020`, na.rm = TRUE),
         sd_GC_19 = sd(`GC/60_2019`, na.rm = TRUE),
         sd_GC_18 = sd(`GC/60_2018`, na.rm = TRUE),
         sd_GC_17 = sd(`GC/60_2017`, na.rm = TRUE),
         sd_GC_16 = sd(`GC/60_2016`, na.rm = TRUE),
         )


playmaker_score <- playmaker3 |> 
  group_by(Player, PlayerID) |> 
  summarise(
    EVA_zscore_2023 = (`EVA/60_2023` - `mean_EVA_23`) / `sd_EVA_23`,
    EVA_zscore_2022 = (`EVA/60_2022` - `mean_EVA_22`) / `sd_EVA_22`,
    EVA_zscore_2021 = (`EVA/60_2021` - `mean_EVA_21`) / `sd_EVA_21`,
    EVA_zscore_2020 = (`EVA/60_2020` - `mean_EVA_20`) / `sd_EVA_20`,
    EVA_zscore_2019 = (`EVA/60_2019` - `mean_EVA_19`) / `sd_EVA_19`,
    EVA_zscore_2018 = (`EVA/60_2018` - `mean_EVA_18`) / `sd_EVA_18`,
    EVA_zscore_2017 = (`EVA/60_2017` - `mean_EVA_17`) / `sd_EVA_17`,
    EVA_zscore_2016 = (`EVA/60_2016` - `mean_EVA_16`) / `sd_EVA_16`,
    
    PPA_zscore_2023 = (`PPA/60_2023` - `mean_PPA_23`) / `sd_PPA_23`,
    PPA_zscore_2022 = (`PPA/60_2022` - `mean_PPA_22`) / `sd_PPA_22`,
    PPA_zscore_2021 = (`PPA/60_2021` - `mean_PPA_21`) / `sd_PPA_21`,
    PPA_zscore_2020 = (`PPA/60_2020` - `mean_PPA_20`) / `sd_PPA_20`,
    PPA_zscore_2019 = (`PPA/60_2019` - `mean_PPA_19`) / `sd_PPA_19`,
    PPA_zscore_2018 = (`PPA/60_2018` - `mean_PPA_18`) / `sd_PPA_18`,
    PPA_zscore_2017 = (`PPA/60_2017` - `mean_PPA_17`) / `sd_PPA_17`,
    PPA_zscore_2016 = (`PPA/60_2016` - `mean_PPA_16`) / `sd_PPA_16`,
    
    GC_zscore_2023 = (`GC/60_2023` - `mean_GC_23`) / `sd_GC_23`,
    GC_zscore_2022 = (`GC/60_2022` - `mean_GC_22`) / `sd_GC_22`,
    GC_zscore_2021 = (`GC/60_2021` - `mean_GC_21`) / `sd_GC_21`,
    GC_zscore_2020 = (`GC/60_2020` - `mean_GC_20`) / `sd_GC_20`,
    GC_zscore_2019 = (`GC/60_2019` - `mean_GC_19`) / `sd_GC_19`,
    GC_zscore_2018 = (`GC/60_2018` - `mean_GC_18`) / `sd_GC_18`,
    GC_zscore_2017 = (`GC/60_2017` - `mean_GC_17`) / `sd_GC_17`,
    GC_zscore_2016 = (`GC/60_2016` - `mean_GC_16`) / `sd_GC_16`
            )


```



```{r}
master_playmaker <- playmaker_score |> 
  select(-PlayerID) |> 
  pivot_longer(-c(Player),
               names_to = c(".value", "Year"), 
               names_sep = "_zscore_") |> 
  group_by(Player, Year) |> 
  mutate(total_zscore = sum(1/2 * EVA, 1/2 * GC)) |> 
  ungroup() |> 
  group_by(Player) |> 
  mutate(mean_zscore = mean(total_zscore, na.rm = TRUE))
```



```{r}
library(ggbeeswarm)
library(ggrepel)


# segment into 4 quartiles for skill level of ability
master_playmaker |> 
  na.omit() |> 
  mutate(Groupings = ifelse(total_zscore <= 0.5 & total_zscore >= -0.5, "Consistent", ifelse(total_zscore >= 0.5, "Good Consistent", "Bad Consistent"))) |> 
  # group_by(Groupings, Year) |> 
  # summarise(mean_EVA = mean(EVA)) |> 
  ggplot(aes(x = Year, y = total_zscore, label = Player)) +
  # geom_point(aes(color = Groupings), alpha = 0.2) +
  geom_beeswarm(alpha = 0.3) +
  geom_hline(yintercept = 0.5, color = "red", linetype = "dashed", size = 2) +
  geom_hline(yintercept = -0.5, color = "red", linetype = "dashed", size = 2) +
  geom_label_repel(size = 2, box.padding = 0.25) +
  scale_x_discrete(expand = c(0.1, 0)) + 
  labs(
    x = "Season Starting Year",
    y = "Z-Score",
    caption = "Data Courtesy of Hockey-Reference"
  ) +
  theme_light()

master_playmaker |> 
  filter(Player == "Sidney Crosby") |> 
  ggplot(aes(x = Year, y = total_zscore)) +
  geom_point() +
  geom_hline(aes(yintercept = mean_zscore)) + 
  labs(
    x = "Season Starting Year",
    y = "Z-Score",
    caption = "Data Courtesy of Hockey-Reference"
  ) +
  theme_light()
```




## Energy Consistency Metric



```{r}
energy <- complete_df_long |> select(Player, PlayerID, starts_with("TOI_"), contains("HIT_"), contains("BLK_"), contains("PIM_"))

energy2 <- energy |> 
  group_by(Player, PlayerID) |> 
  summarise(
    `HIT/60_2023` = (HIT_2023 / TOI_2023) * 60,
    `HIT/60_2022` = (HIT_2022 / TOI_2022) * 60,
    `HIT/60_2021` = (HIT_2021 / TOI_2021) * 60,
    `HIT/60_2020` = (HIT_2020 / TOI_2020) * 60,
    `HIT/60_2019` = (HIT_2019 / TOI_2019) * 60,
    `HIT/60_2018` = (HIT_2018 / TOI_2018) * 60,
    `HIT/60_2017` = (HIT_2017 / TOI_2017) * 60,
    `HIT/60_2016` = (HIT_2016 / TOI_2016) * 60,
    
    `BLK/60_2023` = (BLK_2023 / TOI_2023) * 60,
    `BLK/60_2022` = (BLK_2022 / TOI_2022) * 60,
    `BLK/60_2021` = (BLK_2021 / TOI_2021) * 60,
    `BLK/60_2020` = (BLK_2020 / TOI_2020) * 60,
    `BLK/60_2019` = (BLK_2019 / TOI_2019) * 60,
    `BLK/60_2018` = (BLK_2018 / TOI_2018) * 60,
    `BLK/60_2017` = (BLK_2017 / TOI_2017) * 60,
    `BLK/60_2016` = (BLK_2016 / TOI_2016) * 60,
    
    `PIM/60_2023` = (PIM_2023 / TOI_2023) * 60,
    `PIM/60_2022` = (PIM_2022 / TOI_2022) * 60,
    `PIM/60_2021` = (PIM_2021 / TOI_2021) * 60,
    `PIM/60_2020` = (PIM_2020 / TOI_2020) * 60,
    `PIM/60_2019` = (PIM_2019 / TOI_2019) * 60,
    `PIM/60_2018` = (PIM_2018 / TOI_2018) * 60,
    `PIM/60_2017` = (PIM_2017 / TOI_2017) * 60,
    `PIM/60_2016` = (PIM_2016 / TOI_2016) * 60,
         )

energy3 <- energy2 |> 
  ungroup() |> 
  mutate(mean_HIT_23 = mean(`HIT/60_2023`, na.rm = TRUE),
         mean_HIT_22 = mean(`HIT/60_2022`, na.rm = TRUE),
         mean_HIT_21 = mean(`HIT/60_2021`, na.rm = TRUE),
         mean_HIT_20 = mean(`HIT/60_2020`, na.rm = TRUE),
         mean_HIT_19 = mean(`HIT/60_2019`, na.rm = TRUE),
         mean_HIT_18 = mean(`HIT/60_2018`, na.rm = TRUE),
         mean_HIT_17 = mean(`HIT/60_2017`, na.rm = TRUE),
         mean_HIT_16 = mean(`HIT/60_2016`, na.rm = TRUE),
         
         mean_BLK_23 = mean(`BLK/60_2023`, na.rm = TRUE),
         mean_BLK_22 = mean(`BLK/60_2022`, na.rm = TRUE),
         mean_BLK_21 = mean(`BLK/60_2021`, na.rm = TRUE),
         mean_BLK_20 = mean(`BLK/60_2020`, na.rm = TRUE),
         mean_BLK_19 = mean(`BLK/60_2019`, na.rm = TRUE),
         mean_BLK_18 = mean(`BLK/60_2018`, na.rm = TRUE),
         mean_BLK_17 = mean(`BLK/60_2017`, na.rm = TRUE),
         mean_BLK_16 = mean(`BLK/60_2016`, na.rm = TRUE),
         
         mean_PIM_23 = mean(`PIM/60_2023`, na.rm = TRUE),
         mean_PIM_22 = mean(`PIM/60_2022`, na.rm = TRUE),
         mean_PIM_21 = mean(`PIM/60_2021`, na.rm = TRUE),
         mean_PIM_20 = mean(`PIM/60_2020`, na.rm = TRUE),
         mean_PIM_19 = mean(`PIM/60_2019`, na.rm = TRUE),
         mean_PIM_18 = mean(`PIM/60_2018`, na.rm = TRUE),
         mean_PIM_17 = mean(`PIM/60_2017`, na.rm = TRUE),
         mean_PIM_16 = mean(`PIM/60_2016`, na.rm = TRUE),
         
         
         sd_HIT_23 = sd(`HIT/60_2023`, na.rm = TRUE),
         sd_HIT_22 = sd(`HIT/60_2022`, na.rm = TRUE),
         sd_HIT_21 = sd(`HIT/60_2021`, na.rm = TRUE),
         sd_HIT_20 = sd(`HIT/60_2020`, na.rm = TRUE),
         sd_HIT_19 = sd(`HIT/60_2019`, na.rm = TRUE),
         sd_HIT_18 = sd(`HIT/60_2018`, na.rm = TRUE),
         sd_HIT_17 = sd(`HIT/60_2017`, na.rm = TRUE),
         sd_HIT_16 = sd(`HIT/60_2016`, na.rm = TRUE),
         
         sd_BLK_23 = sd(`BLK/60_2023`, na.rm = TRUE),
         sd_BLK_22 = sd(`BLK/60_2022`, na.rm = TRUE),
         sd_BLK_21 = sd(`BLK/60_2021`, na.rm = TRUE),
         sd_BLK_20 = sd(`BLK/60_2020`, na.rm = TRUE),
         sd_BLK_19 = sd(`BLK/60_2019`, na.rm = TRUE),
         sd_BLK_18 = sd(`BLK/60_2018`, na.rm = TRUE),
         sd_BLK_17 = sd(`BLK/60_2017`, na.rm = TRUE),
         sd_BLK_16 = sd(`BLK/60_2016`, na.rm = TRUE),
         
         sd_PIM_23 = sd(`PIM/60_2023`, na.rm = TRUE),
         sd_PIM_22 = sd(`PIM/60_2022`, na.rm = TRUE),
         sd_PIM_21 = sd(`PIM/60_2021`, na.rm = TRUE),
         sd_PIM_20 = sd(`PIM/60_2020`, na.rm = TRUE),
         sd_PIM_19 = sd(`PIM/60_2019`, na.rm = TRUE),
         sd_PIM_18 = sd(`PIM/60_2018`, na.rm = TRUE),
         sd_PIM_17 = sd(`PIM/60_2017`, na.rm = TRUE),
         sd_PIM_16 = sd(`PIM/60_2016`, na.rm = TRUE),
         )

energy_score <- energy3 |> 
  group_by(Player, PlayerID) |> 
  summarise(
    HIT_zscore_2023 = (`HIT/60_2023` - `mean_HIT_23`) / `sd_HIT_23`,
    HIT_zscore_2022 = (`HIT/60_2022` - `mean_HIT_22`) / `sd_HIT_22`,
    HIT_zscore_2021 = (`HIT/60_2021` - `mean_HIT_21`) / `sd_HIT_21`,
    HIT_zscore_2020 = (`HIT/60_2020` - `mean_HIT_20`) / `sd_HIT_20`,
    HIT_zscore_2019 = (`HIT/60_2019` - `mean_HIT_19`) / `sd_HIT_19`,
    HIT_zscore_2018 = (`HIT/60_2018` - `mean_HIT_18`) / `sd_HIT_18`,
    HIT_zscore_2017 = (`HIT/60_2017` - `mean_HIT_17`) / `sd_HIT_17`,
    HIT_zscore_2016 = (`HIT/60_2016` - `mean_HIT_16`) / `sd_HIT_16`,
    
    BLK_zscore_2023 = (`BLK/60_2023` - `mean_BLK_23`) / `sd_BLK_23`,
    BLK_zscore_2022 = (`BLK/60_2022` - `mean_BLK_22`) / `sd_BLK_22`,
    BLK_zscore_2021 = (`BLK/60_2021` - `mean_BLK_21`) / `sd_BLK_21`,
    BLK_zscore_2020 = (`BLK/60_2020` - `mean_BLK_20`) / `sd_BLK_20`,
    BLK_zscore_2019 = (`BLK/60_2019` - `mean_BLK_19`) / `sd_BLK_19`,
    BLK_zscore_2018 = (`BLK/60_2018` - `mean_BLK_18`) / `sd_BLK_18`,
    BLK_zscore_2017 = (`BLK/60_2017` - `mean_BLK_17`) / `sd_BLK_17`,
    BLK_zscore_2016 = (`BLK/60_2016` - `mean_BLK_16`) / `sd_BLK_16`,
    
    PIM_zscore_2023 = (`PIM/60_2023` - `mean_PIM_23`) / `sd_PIM_23`,
    PIM_zscore_2022 = (`PIM/60_2022` - `mean_PIM_22`) / `sd_PIM_22`,
    PIM_zscore_2021 = (`PIM/60_2021` - `mean_PIM_21`) / `sd_PIM_21`,
    PIM_zscore_2020 = (`PIM/60_2020` - `mean_PIM_20`) / `sd_PIM_20`,
    PIM_zscore_2019 = (`PIM/60_2019` - `mean_PIM_19`) / `sd_PIM_19`,
    PIM_zscore_2018 = (`PIM/60_2018` - `mean_PIM_18`) / `sd_PIM_18`,
    PIM_zscore_2017 = (`PIM/60_2017` - `mean_PIM_17`) / `sd_PIM_17`,
    PIM_zscore_2016 = (`PIM/60_2016` - `mean_PIM_16`) / `sd_PIM_16`
            )
```



```{r}
master_energy <- energy_score |> 
  select(-PlayerID) |> 
  pivot_longer(-c(Player),
               names_to = c(".value", "Year"), 
               names_sep = "_zscore_") |> 
  group_by(Player, Year) |> 
  mutate(total_zscore = sum(2/5 * HIT, 2/5 * BLK, 1/5 * PIM)) |> 
  ungroup() |> 
  group_by(Player) |> 
  mutate(mean_zscore = mean(total_zscore, na.rm = TRUE))
```



```{r}
library(ggbeeswarm)
library(ggrepel)

master_energy |> 
  na.omit() |> 
  mutate(Groupings = ifelse(total_zscore <= 0.5 & total_zscore >= -0.5, "Consistent", ifelse(total_zscore >= 0.5, "Good Consistent", "Bad Consistent"))) |> 
  ggplot(aes(x = Year, y = total_zscore, label = Player)) +
  # geom_point(aes(color = Groupings), alpha = 0.2) +
  geom_beeswarm(alpha = 0.3) +
  # geom_hline(yintercept = 0.5, color = "red", linetype = "dashed", size = 2) +
  # geom_hline(yintercept = -0.5, color = "red", linetype = "dashed", size = 2) +
  geom_label_repel(size = 2, box.padding = 0.5) + 
  labs(
    x = "Season Starting Year",
    y = "Z-Score",
    caption = "Data Courtesy of Hockey-Reference"
  ) + 
  theme_light()

master_energy |> 
  filter(Player == "Ryan Reaves") |> 
  ggplot(aes(x = Year, y = total_zscore)) +
  geom_point() +
  geom_hline(aes(yintercept = mean_zscore)) +
  labs(
    x = "Season Starting Year",
    y = "Z-Score",
    caption = "Data Courtesy of Hockey-Reference"
  ) +
  theme_light()
```





```{r}
# long_misc_df_exp |> 
#   group_by(Player) |> 
#   arrange(desc(total_ptspg_diff)) |> 
#   head(n = 10) |> 
#   ggplot(aes(y = reorder(Player, total_ptspg_diff), x = total_ptspg_diff)) +
#   geom_col(fill = "purple") +
#   geom_label(aes(label = total_ptspg_diff), size = 5) +
#   theme_bw() +
#   labs(
#     x = "Total Points Per Game Difference",
#     y = "Players",
#     caption = "Data courtesy of Hockey-Reference"
#   )
# 
# 
# long_misc_df_exp |> 
#   group_by(Player) |> 
#   arrange(desc(total_ops_diff)) |> 
#   head(n = 10) |> 
#   ggplot(aes(y = reorder(Player, total_ops_diff), x = total_ops_diff)) +
#   geom_col(fill = "purple") +
#   geom_label(aes(label = total_ops_diff), size = 5) +
#   theme_bw() +
#   labs(
#     x = "Total Offensive Point Shares Difference",
#     y = "Players",
#     caption = "Data courtesy of Hockey-Reference"
#   )



# long_misc_df_exp |> 
#   filter(Pos == "F", num_seasons >= 4, consistency_score != Inf) |> 
#   group_by(Player) |> 
#   arrange(desc(consistency_score)) |> 
#   head(n = 10) |> 
#   ggplot(aes(y = reorder(Player, consistency_score), x = consistency_score)) +
#   geom_col(fill = "purple") +
#   geom_label(aes(label = consistency_score), size = 5) +
#   theme_bw() +
#   labs(
#     x = "Preliminary Consistency Score",
#     y = "Players",
#     caption = "Data courtesy of Hockey-Reference"
#   ) +
#   scale_x_continuous(n.breaks = 6)
# 
# 
# long_misc_df_exp |> 
#   filter(Pos == "F", num_seasons >= 6) |> 
#   filter(consistency_score >= 0.5) |> 
#   group_by(Player) |> 
#   arrange(consistency_score) |> 
#   head(n = 10) |> 
#   ggplot(aes(y = reorder(Player, -consistency_score), x = consistency_score)) +
#   geom_col(fill = "purple") +
#   geom_label(aes(label = consistency_score), size = 5) +
#   theme_bw() +
#   labs(
#     x = "Preliminary Consistency Score",
#     y = "Players",
#     caption = "Data courtesy of Hockey-Reference"
#   )
  # scale_x_continuous(limits = c(0, 0.6), n.breaks = 7)

```




























