---
title: "Checkpoint 1 Thoughts"
format: html
editor: visual
author: "Celine, Cole"
---

## Thoughts on measuring consistency

1.  Derive a consistency variable (consistency score) by calculating the differences in performance statistics across seasons; apply different weights by position.

2.  Weight the metrics differently based on importance for each position

3.  Consistency score will have more consistent players near 0 & inconsistent players much greater than 0

4.  Performance metrics must be `per game` or `per 60 minutes` to make it consistent and unbiased among player

Metrics to measure consistency among forwards:

-   Pts/gp

-   +/-

-   Offensive point shares

-   Shooting percentage / shots per game

-   GF/60 (EV, PP, SH)

-   Takeaways? Giveaways?

-   Corsi for %? Fenwick for %?

Metrics to measure consistency among defensemen:

-   +/-

-   Defensive point shares

-   GA/60 (EV, PP, SH)

-   Blocks and hits

-   Takeaways? Giveaways?

Metrics to measure consistency among goalies:

-   Save percentage

-   Goalies against average

-   Quality start percentage

-   Goalie point shares

## Data

[Hockey-reference.com](http://hockey-reference.com/)

### Basic Statistics

```{r}
library(tidyverse)
library(broom)
```

```{r, setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

basic_2023_24 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Basic%20Stats/basic_stats_23_24.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |>
  mutate(start_year = 2023)

basic_2022_23 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Basic%20Stats/basic_stats_22_23.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |>
  mutate(start_year = 2022)

basic_2021_22 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Basic%20Stats/basic_stats_21_22.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |>
  mutate(start_year = 2021)

basic_2020_21 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Basic%20Stats/basic_stats_20_21.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |>
  mutate(start_year = 2020)

basic_2019_20 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Basic%20Stats/basic_stats_19_20.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |>
  mutate(start_year = 2019)

basic_2018_19 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Basic%20Stats/basic_stats_18_19.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |>
  mutate(start_year = 2018)

basic_2017_18 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Basic%20Stats/basic_stats_17_18.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |>
  mutate(start_year = 2017)

basic_2016_17 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Basic%20Stats/basic_stats_16_17.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |>
  mutate(start_year = 2016)

```

```{r}
# combined basic stats
basic_df <- rbind(basic_2023_24, basic_2022_23, basic_2021_22, basic_2020_21, basic_2021_22, 
                     basic_2020_21, basic_2019_20, basic_2018_19, basic_2017_18, basic_2016_17)
```

```{r, setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

long_basic_df <- basic_df |> 
  pivot_wider(id_cols = c(Player, PlayerID), 
              values_from = c(G, A, PTS, `+/-`, `S%`, PS), 
              names_from = start_year)

long_basic_df <- long_basic_df %>% 
  unnest() %>% 
  replace(. == "NULL", NA)
  # mutate(across(G_2023:PS_2016, ~replace_na(.x, 0)))

long_basic_df <- long_basic_df |> 
  unique()
```

### Miscellaneous Statistics

```{r, setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

misc_2023_24 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Misc%20Stats/misc_stats_23_24.csv") |> 
  mutate(start_year = 2023)

misc_2022_23 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Misc%20Stats/misc_stats_22_23.csv") |>
  mutate(start_year = 2022)

misc_2021_22 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Misc%20Stats/misc_stats_21_22.csv") |> 
  mutate(start_year = 2021)

misc_2020_21 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Misc%20Stats/misc_stats_20_21.csv") |> 
  mutate(start_year = 2020)

misc_2019_20 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Misc%20Stats/misc_stats_19_20.csv") |> 
  mutate(start_year = 2019)

misc_2018_19 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Misc%20Stats/misc_stats_18_19.csv") |> 
  mutate(start_year = 2018)

misc_2017_18 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Misc%20Stats/misc_stats_17_18.csv") |> 
  mutate(start_year = 2017)

misc_2016_17 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Misc%20Stats/misc_stats_16_17.csv") |> 
  mutate(start_year = 2016)
```

```{r}
misc_df <- rbind(misc_2023_24, misc_2022_23, misc_2021_22, misc_2020_21, misc_2019_20,
                 misc_2018_19, misc_2017_18, misc_2016_17)
```

```{r, setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

long_misc_df <- misc_df |> 
  filter(Pos != "G") |> 
  pivot_wider(id_cols = c(Player), 
              values_from = c(`G/GP`, `A/GP`, `PTS/GP`, `GC/GP`, `S/GP`, `OPS`, `DPS`, `PS`), 
              names_from = start_year)

long_misc_df <- long_misc_df %>% 
  unnest() %>% 
  replace(. == "NULL", NA)
  # mutate(across(`G/GP_2023`:`PS_2016`, ~replace_na(.x, 0)))

long_misc_df_exp <- long_misc_df |>
  group_by(Player) |> 
  summarize(ptspg_diff_23_22 = abs(`PTS/GP_2023` - `PTS/GP_2022`),
         ptspg_diff_22_21 = abs(`PTS/GP_2022` - `PTS/GP_2021`),
         ptspg_diff_21_20 = abs(`PTS/GP_2021` - `PTS/GP_2020`),
         ptspg_diff_20_19 = abs(`PTS/GP_2020` - `PTS/GP_2019`),
         ptspg_diff_19_18 = abs(`PTS/GP_2019` - `PTS/GP_2018`),
         ptspg_diff_18_17 = abs(`PTS/GP_2018` - `PTS/GP_2017`),
         ptspg_diff_17_16 = abs(`PTS/GP_2017` - `PTS/GP_2016`),
         ops_diff_23_22 = abs(OPS_2023 - OPS_2022),
         ops_diff_22_21 = abs(OPS_2022 - OPS_2021),
         ops_diff_21_20 = abs(OPS_2021 - OPS_2020),
         ops_diff_20_19 = abs(OPS_2020 - OPS_2019),
         ops_diff_19_18 = abs(OPS_2019 - OPS_2018),
         ops_diff_18_17 = abs(OPS_2018 - OPS_2017),
         ops_diff_17_16 = abs(OPS_2017 - OPS_2016)
         )
```

### Advanced Statistics

```{r, setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

advanced_2023_24 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Advanced%20Stats/advanced_stats_23_24.csv") |> 
  mutate(start_year = 2023)

advanced_2022_23 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Advanced%20Stats/advanced_stats_22_23.csv") |> 
  mutate(start_year = 2022)

advanced_2021_22 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Advanced%20Stats/advanced_stats_21_22.csv") |> 
  mutate(start_year = 2021)

advanced_2020_21 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Advanced%20Stats/advanced_stats_20_21.csv") |> 
  mutate(start_year = 2020)

advanced_2019_20 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Advanced%20Stats/advanced_stats_19_20.csv") |> 
  mutate(start_year = 2019)

advanced_2018_19 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Advanced%20Stats/advanced_stats_18_19.csv") |> 
  mutate(start_year = 2018)

advanced_2017_18 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Advanced%20Stats/advanced_stats_17_18.csv") |> 
  mutate(start_year = 2017)

advanced_2016_17 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Advanced%20Stats/advanced_stats_16_17.csv") |> 
  mutate(start_year = 2016)
```

```{r}
advanced_df <- rbind(advanced_2023_24, advanced_2022_23, advanced_2021_22, advanced_2020_21,
                     advanced_2019_20, advanced_2018_19, advanced_2017_18, advanced_2016_17)
```

## EDAs

```{r}
library(ggplot2)
```

### EDAs using Basic Stats

### EDAs using Miscellaneous Stats

```{r}
long_misc_df_exp |> 
  group_by(Player) |> 
  summarise(total_ptspg_diff = sum(ptspg_diff_23_22, ptspg_diff_22_21, ptspg_diff_21_20, ptspg_diff_20_19, ptspg_diff_19_18, ptspg_diff_18_17, ptspg_diff_17_16)) |> 
  arrange(desc(total_ptspg_diff)) |> 
  head(n = 10) |> 
  ggplot(aes(y = reorder(Player, total_ptspg_diff), x = total_ptspg_diff)) +
  geom_col(fill = "purple") +
  geom_label(aes(label = total_ptspg_diff), size = 5) +
  theme_bw() +
  labs(
    x = "Total Points Per Game Difference",
    y = "Players",
    caption = "Data courtesy of Hockey-Reference"
  )


long_misc_df_exp |> 
  group_by(Player) |> 
  summarise(total_ops_diff = sum(ops_diff_23_22, ops_diff_22_21, ops_diff_21_20, ops_diff_20_19, ops_diff_19_18, ops_diff_18_17, ops_diff_17_16)) |> 
  arrange(desc(total_ops_diff)) |> 
  head(n = 10) |> 
  ggplot(aes(y = reorder(Player, total_ops_diff), x = total_ops_diff)) +
  geom_col(fill = "purple") +
  geom_label(aes(label = total_ops_diff), size = 5) +
  theme_bw() +
  labs(
    x = "Total Points Per Game Difference",
    y = "Players",
    caption = "Data courtesy of Hockey-Reference"
  )
```

```{r}
# shooting percentage difference

long_misc_df |>
  group_by(Player) |> 
  summarize(spg_diff_23_22 = abs(`S/GP_2023` - `S/GP_2022`),
            spg_diff_22_21 = abs(`S/GP_2022` - `S/GP_2021`),
            spg_diff_21_20 = abs(`S/GP_2021` - `S/GP_2020`),
            spg_diff_20_19 = abs(`S/GP_2020` - `S/GP_2019`),
            spg_diff_19_18 = abs(`S/GP_2019` - `S/GP_2018`),
            spg_diff_18_17 = abs(`S/GP_2018` - `S/GP_2017`),
            spg_diff_17_16 = abs(`S/GP_2017` - `S/GP_2016`),
            total_spg_diff = sum(spg_diff_23_22, spg_diff_22_21, spg_diff_21_20, spg_diff_20_19,
                                 spg_diff_19_18, spg_diff_18_17, spg_diff_17_16)) |> 
  arrange(desc(total_spg_diff)) |> 
  head(n = 10) |> 
  ggplot(aes(y = reorder(Player, total_spg_diff), x = total_spg_diff)) +
  geom_col(color = "grey", fill = "lightblue") +
  geom_label(aes(label = total_spg_diff), size = 3) +
  labs(x = "Difference in Total Shooting Percentage Per Game", y = "Players") +
  theme_minimal()
```

### EDAs using Advanced Stats

```{r}
advanced_df |> 
  select("Player", "Pos", "CF%", "FF%", "oiSH%", "oiSV%", "oZS%", "dZS%", "start_year") |> 
  filter(Player == "Jake Guentzel") |> 
  pivot_longer(cols = c("CF%", "FF%", "oiSH%", "oiSV%", "oZS%", "dZS%"),
               names_to = "Stats",
               values_to = "Value") |> 
  ggplot(aes(x = start_year, y = Value, color = Stats)) +
  geom_point(alpha = 0.6) +
  geom_smooth(se = FALSE) +
  labs(title = "Metrics Over the Years for Evgeni Malkin",
       x = "Year",
       y = "Stats Value",
       color = "Stats") +
  theme_minimal() +
  scale_x_continuous(n.breaks = 8)
```
