---
title: "Checkpoint 1 Thoughts"
format: html
editor: visual
author: "Celine, Cole"
---

## Thoughts on measuring consistency

1.  Derive a consistency variable (consistency score) by calculating the differences in performance statistics across seasons; apply different weights by position.

2.  Weight the metrics differently based on importance for each position

3.  Consistency score will have more consistent players near 0 & inconsistent players much greater than 0

4.  Performance metrics must be `per game` or `per 60 minutes` to make it consistent and unbiased among player

## Data

### Basic Statistics

### Advanced Statistics

```{r}
Stats_AD23.24 <- read.csv("~/Desktop/Penguins Capstone Project/Advanced Stats/Stats_AD23-24.csv")

Stats_AD22.23 <- read.csv("~/Desktop/Penguins Capstone Project/Advanced Stats/Stats_AD22-23.csv")

Stats_AD21.22 <- read.csv("~/Desktop/Penguins Capstone Project/Advanced Stats/Stats_AD21-22.csv")

Stats_AD20.21 <- read.csv("~/Desktop/Penguins Capstone Project/Advanced Stats/Stats_AD20-21.csv")

Stats_AD19.20 <- read.csv("~/Desktop/Penguins Capstone Project/Advanced Stats/Stats_AD19-20.csv")

Stats_AD18.19 <- read.csv("~/Desktop/Penguins Capstone Project/Advanced Stats/Stats_AD18-19.csv")

Stats_AD17.18 <- read.csv("~/Desktop/Penguins Capstone Project/Advanced Stats/Stats_AD17-18.csv")

Stats_AD16.17 <- read.csv("~/Desktop/Penguins Capstone Project/Advanced Stats/Stats_AD16-17.csv")

combined_df <- rbind(Stats_AD23.24, Stats_AD22.23, Stats_AD21.22, Stats_AD20.21, Stats_AD19.20, 
                     Stats_AD18.19, Stats_AD17.18, Stats_AD16.17)
```

## EDAs

### EDAs using Basic Stats

### EDAs using Advanced Stats

```{r}
library(ggplot2)
library(tidyverse)
```

```{r}
combined_df |> 
  select("Player", "Pos", "CF.", "FF.", "oiSH.", "oiSV.", "oZS.", "dZS.", "Year") |> 
  filter(Player == "Jake Guentzel") |> 
  pivot_longer(cols = c(CF., FF., oiSH., oiSV., oZS., dZS.),
               names_to = "Metric",
               values_to = "Value") |> 
  ggplot(aes(x = Year, y = Value, color = Metric)) +
  geom_point(alpha = 0.6) +
  geom_smooth(se = FALSE) +
  labs(title = "Metrics Over the Years for Evgeni Malkin",
       x = "Year",
       y = "Metric Value",
       color = "Metric") +
  theme_minimal()
```
