---
title: "Sketching Consistency"
author: "Chisunta M Chikwamu"
format: html
editor: visual
---

## Consistency in Hockey

How do we define a consistent player?

**Initial thoughts to be investigated;**

-   Consistent goalies, defense men, forward men or what combinations of consistency make an optimal outcome

-   Are these combinations team specific? How do we find optimal combinations.

    ### Understanding Hockey Data or what kind of data to use:

    -   Play by play

    -   Team data or player data

    -   Metrics to look at:

        -   means, variances within a season of valuable performance
        -   Need to define valuable performance based on player type i.e goalie, forward or defense man

## Obtaining Data

```{r}
install.packages("rvest")
```

```{r}
library(rvest)
library(tidyverse)
nhl_url <- "https://www.hockey-reference.com/teams/PIT/"
nhl_url |> 
  read_html()
```

```{r}
nhl_tbl_2024 <- nhl_url |> 
  read_html() |> 
  html_element(css = "#div_skaters") |> 
  html_table()
nhl_tbl_2024
```

```{r}
nhl_tbl_2024 <- nhl_url |> 
  read_html() |> 
  html_element(css = "#div_skaters") |> 
  html_table()
nhl_tbl_2024


```

Distribution of Positions:

```{r}
# Assign the first row as column names
colnames(nhl_tbl_2024) <- nhl_tbl_2024[1, ]
nhl_tbl_2024 <- nhl_tbl_2024[-1, ]

```

```{r}
nhl_tbl_2024 |>
  select(Pos)|> 
  group_by(Pos)|>
  count()
```

```{r}
basic_2023_24 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Basic%20Stats/basic_stats_23_24.csv") |> 
  rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |>
  mutate(start_year = 2023)
```

```{r}
basic_2023_24
```

```{r}
#filtering by position and selecting some metrics that maybe used to calculate and assess quality of defensemen
defensemen <- basic_2023_24|>
  filter(Pos == "D")|>
  select(Player, Age, GP, G, A, PTS,PS, S, TOI, BLK, HIT)
glimpse(defensemen)
```


##Some very uneccessary EDA to just get a sense of the distribution of variables
```{r}
theme_set(theme_light())
# Plot the distribution of each variable
variables <- c("Age", "GP", "G", "A", "PTS", "PS", "S", "TOI", "BLK", "HIT")

# Create plots for each variable
for (var in variables) {
  p <- ggplot(defensemen, aes_string(x = var)) +
    geom_histogram(binwidth = 1, fill = "lightblue") +
    theme_minimal() +
    labs(title = paste("Distribution of", var), x = var, y = "Frequency")
  
  print(p)
}
```


```{r}
# Calculate Blocks per Game (BPG) and Blocks per Minute (BPM)
defensemen <- defensemen |>
  mutate(
    BPG = BLK / GP,
    BPM = BLK / TOI
  )
# new defensemen with BPG and BPM
glimpse(defensemen)
```

```{r}

# Visualize Blocks per Game (BPG) and Blocks per Minute (BPM)

# Blocks per Game
ggplot(defensemen, aes(x = BPG)) +
  geom_histogram(binwidth = 0.1, fill = "blue", color = "black") +
  theme_minimal() +
  labs(title = "Distribution of Blocks per Game (BPG) for Defensemen", x = "Blocks per Game", y = "Frequency")

# Blocks per Minute
ggplot(defensemen, aes(x = BPM)) +
  geom_histogram(binwidth = 0.001, fill = "green", color = "black") +
  theme_minimal() +
  labs(title = "Distribution of Blocks per Minute (BPM) for Defensemen", x = "Blocks per Minute", y = "Frequency")

# Scatter plot to visualize consistency
ggplot(defensemen, aes(x = GP, y = BPG)) +
  geom_point(aes(size = TOI, color = Player)) +
  theme_minimal() +
  labs(title = "Blocks per Game vs. Games Played", x = "Games Played (GP)", y = "Blocks per Game (BPG)")

ggplot(defensemen, aes(x = TOI, y = BPM)) +
  geom_point(aes(size = GP, color = Player)) +
  theme_minimal() +
  labs(title = "Blocks per Minute vs. Time on Ice", x = "Time on Ice (TOI)", y = "Blocks per Minute (BPM)")


```


##Data

### Basic Statistics

```{r}
library(tidyverse)
library(broom)

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

basic_2023_24 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Basic%20Stats/basic_stats_23_24.csv") |>
rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |>
mutate(start_year = 2023)

basic_2022_23 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Basic%20Stats/basic_stats_22_23.csv") |>
rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |>
mutate(start_year = 2022)

basic_2021_22 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Basic%20Stats/basic_stats_21_22.csv") |>
rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |>
mutate(start_year = 2021)

basic_2020_21 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Basic%20Stats/basic_stats_20_21.csv") |>
rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |>
mutate(start_year = 2020)

basic_2019_20 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Basic%20Stats/basic_stats_19_20.csv") |>
rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |>
mutate(start_year = 2019)

basic_2018_19 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Basic%20Stats/basic_stats_18_19.csv") |>
rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |>
mutate(start_year = 2018)

basic_2017_18 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Basic%20Stats/basic_stats_17_18.csv") |>
rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |>
mutate(start_year = 2017)

basic_2016_17 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Basic%20Stats/basic_stats_16_17.csv") |>
rename(EVG = EV...13, PPG = PP...14, SHG = SH...15, EVA = EV...17, PPA = PP...18, SHA = SH...19) |>
mutate(start_year = 2016)
# combined basic stats
basic_df <- rbind(basic_2023_24, basic_2022_23, basic_2021_22, basic_2020_21, basic_2021_22, 
                     basic_2020_21, basic_2019_20, basic_2018_19, basic_2017_18, basic_2016_17)
```



```{r}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
long_basic_df <- basic_df |> 
  pivot_wider(id_cols = c(Player, PlayerID), 
              values_from = c(G, A, PTS, `+/-`, `S%`, PS), 
              names_from = start_year)

long_basic_df <- long_basic_df |>
  unnest() |> 
  mutate(across(everything(), ~ifelse(. == "NULL", NA, .)))

long_basic_df <- long_basic_df |> 
  unique()

```

### Miscellaneous Statistics

```{r, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

misc_2023_24 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Misc%20Stats/misc_stats_23_24.csv") |>
mutate(start_year = 2023)

misc_2022_23 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Misc%20Stats/misc_stats_22_23.csv") |>
mutate(start_year = 2022)

misc_2021_22 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Misc%20Stats/misc_stats_21_22.csv") |>
mutate(start_year = 2021)

misc_2020_21 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Misc%20Stats/misc_stats_20_21.csv") |>
mutate(start_year = 2020)

misc_2019_20 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Misc%20Stats/misc_stats_19_20.csv") |>
mutate(start_year = 2019)

misc_2018_19 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Misc%20Stats/misc_stats_18_19.csv") |>
mutate(start_year = 2018)

misc_2017_18 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Misc%20Stats/misc_stats_17_18.csv") |>
mutate(start_year = 2017)

misc_2016_17 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Misc%20Stats/misc_stats_16_17.csv") |>
mutate(start_year = 2016)

misc_df <- rbind(misc_2023_24, misc_2022_23, misc_2021_22, misc_2020_21, misc_2019_20,  misc_2018_19, misc_2017_18, misc_2016_17)



```

## Some metrics based on Mis stats

```{r} knitr::opts_chunk$set(warning = FALSE, message = FALSE)}

long_misc_df <- misc_df |>
filter(Pos != "G") |>
pivot_wider(id_cols = c(Player),
values_from = c(`G/GP`, `A/GP`, `PTS/GP`, `GC/GP`, `S/GP`, `OPS`, `DPS`, `PS`),
names_from = start_year)

long_misc_df <- long_misc_df |>
unnest() |>
replace(. == "NULL", NA)
# mutate(across(`G/GP_2023`:`PS_2016`, ~replace_na(.x, 0)))

long_misc_df_exp <- long_misc_df |>
group_by(Player) |>
summarize(ptspg_diff_23_22 = abs(`PTS/GP_2023` - `PTS/GP_2022`),
ptspg_diff_22_21 = abs(`PTS/GP_2022` - `PTS/GP_2021`),
ptspg_diff_21_20 = abs(`PTS/GP_2021` - `PTS/GP_2020`),
ptspg_diff_20_19 = abs(`PTS/GP_2020` - `PTS/GP_2019`),
ptspg_diff_19_18 = abs(`PTS/GP_2019` - `PTS/GP_2018`),
ptspg_diff_18_17 = abs(`PTS/GP_2018` - `PTS/GP_2017`),
ptspg_diff_17_16 = abs(`PTS/GP_2017` - `PTS/GP_2016`),
ops_diff_23_22 = abs(OPS_2023 - OPS_2022),
ops_diff_22_21 = abs(OPS_2022 - OPS_2021),
ops_diff_21_20 = abs(OPS_2021 - OPS_2020),
ops_diff_20_19 = abs(OPS_2020 - OPS_2019),
ops_diff_19_18 = abs(OPS_2019 - OPS_2018),
ops_diff_18_17 = abs(OPS_2018 - OPS_2017),
ps_diff_17_16 = abs(OPS_2017 - OPS_2016))
```

### Advanced Statistics

```{r, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

advanced_2023_24 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Advanced%20Stats/advanced_stats_23_24.csv") |>
mutate(start_year = 2023)

advanced_2022_23 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Advanced%20Stats/advanced_stats_22_23.csv") |>
mutate(start_year = 2022)

advanced_2021_22 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Advanced%20Stats/advanced_stats_21_22.csv") |>
mutate(start_year = 2021)

advanced_2020_21 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Advanced%20Stats/advanced_stats_20_21.csv") |>
mutate(start_year = 2020)

advanced_2019_20 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Advanced%20Stats/advanced_stats_19_20.csv") |>
mutate(start_year = 2019)

advanced_2018_19 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Advanced%20Stats/advanced_stats_18_19.csv") |>
mutate(start_year = 2018)

advanced_2017_18 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Advanced%20Stats/advanced_stats_17_18.csv") |>
mutate(start_year = 2017)

advanced_2016_17 <- read_csv("https://raw.githubusercontent.com/CelineS138/Penguins-Capstone-Project/main/Data/Advanced%20Stats/advanced_stats_16_17.csv") |>
mutate(start_year = 2016)
advanced_df <- rbind(advanced_2023_24, advanced_2022_23, advanced_2021_22, advanced_2020_21, advanced_2019_20, advanced_2018_19, advanced_2017_18, advanced_2016_17)
```

 ## Analysis
 
```{r}
library(dplyr)

# Calculate league averages for points per game and time on ice per game each year
league_avg_stats <- long_basic_df |>
  group_by(start_year)|>
  summarise(
    league_avg_pts_per_gp = mean(PTS/GP, na.rm = TRUE),
    league_avg_toi_per_gp = mean(TOI/GP, na.rm = TRUE)
  )

# Join the averages back to the original data and calculate normalized metrics
basic_df <- long_basic_df|>
  left_join(league_avg_stats, by = "start_year") |>
  mutate(
    norm_pts_per_gp = (PTS/GP) / league_avg_pts_per_gp,
    norm_toi_per_gp = (TOI/GP) / league_avg_toi_per_gp
  )

# Display the resulting dataframe
print(basic_df)

```
 

 
 